### 2024-08-14 paper/paper.tex
Add empty paper dir

First, we seek to generate a new Reynolds-averaged Navier--Stokes (RANS)
turbulence model that accurately predicts a turbulent DNS flow field.
We do this by creating new PDE terms based on the mean velocity and pressure
fields,
then use a simple regression technique to solve for the coefficients of
these terms.

Then, we simulate the flow field with this new model, showing how it is
much more accurate than a laminar model, i.e., one with no turbulence
parameterization.